\chapter{Conclusion}
\label{cha:7}

\subsection{Remarks}

Simpler, frequency-based methods such as BoW and TF-IDF with logistic regression already provides decent overall result. Transfomer-based approach 

Given that the dataset imbalance is solved, it is assumed that BERT and MAGPIE methods should 

While we can seemingly just take the outlet and use this information to classify media bias...

Overall, including outlet information increase the performance in all methods, some more significant than others.

Generally, I argue that it is better to have a model that generalise well over article content rather than relying on outlet and metadata information to classify media bias in articles.

Furthermore, we have not tested the model with unseen data containing unseen outlets...3

The whole dataset is build by Ad Fontes, articles are picked by Ad Fontes, annotations are done by people from Ad Fontes, therefore there is an underlying bias that exist original to the dataset, affecting models trained with this dataset.

It is really difficult to compare this result with other works on media bias classification as there is no comparable dataset particularly for article-level media bias classification.

\subsection{Future Work}

The methods need to be reevaluated with a bigger dataset and particularly much more examples of biased articles, as the current dataset is highly-imbalanced.

Classifying media bias right now, in this way, does not provide much explanaibility. Ultimately, we would want a model to also output a reasoning behind the classification.

Implement a more global methods using graph-based approach to encode multiple articles and define relationships between them. This would allow for a bigger and more detailed representation of articles, allowing the model to generalise according to combined features of the corresponding articles.

Carefully tune the test set to enable more representational articles and pattern, include unseen outlets.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: 
